---
title: "ZooniverseProcessing"
author: "Taiki Sakai"
date: "2022-12-14"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message=FALSE)
```

```{r}
library(rjson)
library(dplyr)
library(here)
library(ggplot2)
library(viridis)
library(caret)
library(likert)
```

```{r}
classCsv <- here('data', 'Zooniverse Labels', 'ocean-voices-classifications.csv')
classData <- read.csv(classCsv, stringsAsFactors = FALSE)
classData$created_at <- as.POSIXct(classData$created_at, format='%Y-%m-%d %H:%M:%S UTC', tz='UTC')

labelCsv <- here('data','Zooniverse Labels', 'CCES_010_BetaTest150subjects_labeled.csv')
labelData <- read.csv(labelCsv, stringsAsFactors = FALSE)
```

We only want to look at data from most recent workflows 8.15 and 17.2

```{r}
classData <- filter(classData, workflow_version %in% c('8.15', '17.2'))
```

Check number of labels selected for each annoation

```{r}
nLabels <- sapply(classData$annotations, function(x) {
  length(fromJSON(x)[[1]]$value)
})
names(nLabels) <- NULL
table(nLabels)
```

"Multiple selections OK" was included by default, so we want to re-work
these multi-label cases.

```{r}
parseValues <- function(x) {
  # single selections are fine, also renaming one caes to another
  if(length(x) == 1 &&
     x != 'Humpback whales and ship noise') {
    return(x)
  }
  # if you selected nothing & something thats confusing, so drop it
  if('None of the above' %in% x) {
    return(NA)
  }
  # otherwise convert to both
  'Humpback whales and ships'
}
classData$label <- sapply(classData$annotations, function(x) {
  parseValues(fromJSON(x)[[1]]$value)
})
classData$label <- factor(classData$label)
table(classData$label, classData$workflow_version)
classData$image <- sapply(classData$subject_data, function(x) {
  fromJSON(x)[[1]]$subject_id
})
```

We want to look at only the retired images, images that are not yet
retired will have `NULL` for the retired portion of that JSON

```{r}
classData$retired <- sapply(classData$subject_data, function(x) {
  json <- fromJSON(x)[[1]]
  length(json$retired) > 0
})
table(classData$retired)
classData <- classData[classData$retired, ]
```

Now we want to look at the distribution of labels for each image.

```{r}
imgList <- lapply(split(classData, classData$subject_ids), function(x) {
  x[c('user_id', 'workflow_version', 'label', 'subject_ids', 'created_at', 'image')]
})
# if not NULL, num limits the number of samples to use
labelDist <- function(x, num=NULL) {
  if(!is.null(num)) {
    num <- min(num, nrow(x))
    x <- x[1:num, ]
  }
  tab <- table(x$label)
  n <- sum(tab)
  tab <- tab / n
  whichMax <- tab == max(tab)
  if(sum(whichMax) > 1) {
    label <- paste0(sum(whichMax), ' tied')
  } else {
    label <- names(tab)[whichMax]
  }
  max <- tab[whichMax][1]
  result <- as.list(tab)
  result$n <- n
  result$max <- max
  result$maxLabel <- label
  result$image <- x$image[1]
  result
}
maxLabs <- lapply(imgList, function(x) {
  labelDist(x, num=NULL)
})
maxLabs <- bind_rows(maxLabs, .id='subject')
```

The `maxLabs` dataframe has one row for every image. Each row shows the
proportion of votes for each label, the number of votes, the max
proportion value, and the label for the max value. In case of ties for
max proportion these are listed as "N tied".

```{r}
str(maxLabs)
```

Plotting distribution of most-confident class for each image. Ideally
these would all be 1 if all reviewers agreed on the label

```{r}
ggplot(maxLabs, aes(x=max, fill=maxLabel)) +
  geom_histogram(binwidth=.05)
```

Compare labels for each image with expert label.

```{r}
userLabs <- select(labelData,subject_id,Expert.Label) %>%
  rename(image=subject_id)%>%
  merge(classData) 

cm<-confusionMatrix(as.factor(userLabs$Expert.Label), as.factor(userLabs$label), positive = "True")

cm_d <- as.data.frame(cm$table) # extract the confusion matrix values as data.frame
cm_st <-data.frame(cm$overall) # confusion matrix statistics as data.frame
cm_st$cm.overall <- round(cm_st$cm.overall,2) # round the values
cm_d$diag <- cm_d$Prediction == cm_d$Reference # Get the Diagonal
cm_d$ndiag <- cm_d$Prediction != cm_d$Reference # Off Diagonal     
cm_d[cm_d == 0] <- NA # Replace 0 with NA for white tiles
cm_d$Reference <-  reverse.levels(cm_d$Reference) # diagonal starts at top left
cm_d$ref_freq <- cm_d$Freq * ifelse(is.na(cm_d$diag),-1,1)

ggplot(data = cm_d, aes(x = Prediction , y =  Reference, fill = Freq))+
  scale_x_discrete(position = "top") +
  geom_tile(data = cm_d,aes(fill = ref_freq)) +
  scale_fill_viridis() +
  geom_text(aes(label = Freq), color = 'black', size = 3)+
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        legend.position = "none",
        panel.border = element_blank(),
        plot.background = element_blank(),
        axis.line = element_blank(),
        axis.text.x = element_text(angle = 45, vjust = 1, hjust=0)
        )

```
