---
title: "ZooniverseProcessing"
author: "Taiki Sakai"
date: "2022-12-14"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(rjson)
library(dplyr)
library(here)
library(ggplot2)
```

```{r}
classCsv <- here('data', 'Zooniverse Labels', 'ocean-voices-classifications.csv')
classData <- read.csv(classCsv, stringsAsFactors = FALSE)
classData$created_at <- as.POSIXct(classData$created_at, format='%Y-%m-%d %H:%M:%S UTC', tz='UTC')
```

We only want to look at data from most recent workflows 8.15 and 17.2

```{r}
classData <- filter(classData, workflow_version %in% c('8.15', '17.2'))
```

Check number of labels selected for each annoation

```{r}
nLabels <- sapply(classData$annotations, function(x) {
  length(fromJSON(x)[[1]]$value)
})
names(nLabels) <- NULL
table(nLabels)
```

"Multiple selections OK" was included by default, so we want to re-work these 
multi-label cases.

```{r}
parseValues <- function(x) {
  # single selections are fine, also renaming one caes to another
  if(length(x) == 1 &&
     x != 'Humpback whales and ship noise') {
    return(x)
  }
  # if you selected nothing & something thats confusing, so drop it
  if('None of the above' %in% x) {
    return(NA)
  }
  # otherwise convert to both
  'Humpback whales and ships'
}
classData$label <- sapply(classData$annotations, function(x) {
  parseValues(fromJSON(x)[[1]]$value)
})
classData$label <- factor(classData$label)
table(classData$label, classData$workflow_version)
classData$image <- sapply(classData$subject_data, function(x) {
  fromJSON(x)[[1]]$subject_id
})
```

We want to look at only the retired images, images that are not yet retired
will have `NULL` for the retired portion of that JSON

```{r}
classData$retired <- sapply(classData$subject_data, function(x) {
  json <- fromJSON(x)[[1]]
  length(json$retired) > 0
})
table(classData$retired)
classData <- classData[classData$retired, ]
```

Now we want to look at the distribution of labels for each image. 

```{r}
imgList <- lapply(split(classData, classData$subject_ids), function(x) {
  x[c('user_id', 'workflow_version', 'label', 'subject_ids', 'created_at', 'image')]
})
# if not NULL, num limits the number of samples to use
labelDist <- function(x, num=NULL) {
  if(!is.null(num)) {
    num <- min(num, nrow(x))
    x <- x[1:num, ]
  }
  tab <- table(x$label)
  n <- sum(tab)
  tab <- tab / n
  whichMax <- tab == max(tab)
  if(sum(whichMax) > 1) {
    label <- paste0(sum(whichMax), ' tied')
  } else {
    label <- names(tab)[whichMax]
  }
  max <- tab[whichMax][1]
  result <- as.list(tab)
  result$n <- n
  result$max <- max
  result$maxLabel <- label
  result$image <- x$image[1]
  result
}
maxLabs <- lapply(imgList, function(x) {
  labelDist(x, num=NULL)
})
maxLabs <- bind_rows(maxLabs, .id='subject')
```

The `maxLabs` dataframe has one row for every image. Each row shows the proportion of 
votes for each label, the number of votes, the max proportion value, and the
label for the max value. In case of ties for max proportion these are listed as
"N tied".

```{r}
str(maxLabs)
```

Plotting distribution of most-confident class for each image. Ideally these
would all be 1 if all reviewers agreed on the label

```{r}
ggplot(maxLabs, aes(x=max, fill=maxLabel)) +
  geom_histogram(binwidth=.05)
```